Theories of deliberated judgment

### 2022-08-11 simple.tex
Rm a lot of stuff

\subsection{Old definition}
\label{sec:descrdef}

Here is an example of a theory that does not use an argumentation strategy, but simply plays an argument that it claims is decisive for everybody.
\begin{example}[Constant argumentation]
	Given the normative theory of \cref{ex:twolott}, define ${\gleadsto} = \set{(\ar_0, D > P)}$ and $\dg = \emptyset$. This defines a descriptive theory of depth $1$ for that normative theory.
\end{example}

\begin{example}[A theory for $P$]
	\label{ex:twolottstrings}
	Define a normative theory as in \cref{ex:twolott}, with an extended set of arguments: define $\allargs$ as the set of all strings. (This correspondingly changes $\allhist$ and thus $\ileadsto$, which must be extended to its new domain.) Define $\ar_0$ as in \cref{ex:twolott}.
	Define $\ar_\text{for$P$}$ as the text “We recommend to adopt $P$: it has a much higher probability of winning.” 
	Define ${\gleadsto} = \set{(\ar_\text{for$P$}, P > D)}$.
	In order to define a counter-argument to $\ar_0$, given a real number $r \in \R$, define $\ar^r_\text{non-linear}$ as the text “$D$ is better than $P$ in terms of expected gain, but it is in general absurd to reason in terms of expected gain. For example, consider a choice between winning $r$ dollars with probability $0.8$ and winning $2r$ dollars with probability $0.4$. They have the same expected gain, but the first one is clearly preferrable. Therefore, we recommend to rather pick $P$ as it presents a much better chance of earning money”. (The value of $r$ and $2r$ is to be used in the relevant places in  the text.)
	Define $\dg = \set{(\ar_\text{for$P$}, \hist, \ar^{10^6}_\text{non-linear}) \suchthat \hist \in \allhist \land \ar_0 \in \hist}$. This strategy anticipates a possible attack using $\ar_0$ against $\ar_\text{for$P$}$, and prepares $\ar^{10^6}_\text{non-linear}$ as a counter-argument.
	This defines a descriptive theory of depth $2$.
\end{example}

The following less trivial example of a descriptive theory illustrates a theory that claims to be able to answer an infinite number of arguments. This also shows that the bounded depth constraint does not imply finiteness of the relation $\dg$.
\begin{example}[Infinite $\dg$]
	\label{ex:infinite}
	Define a normative theory and the arguments $\ar_0$ and $\ar^r_\text{non-linear}$ for any $r \in \R$ as in \cref{ex:twolottstrings}.
	Define ${\gleadsto} = \set{(\ar_0, D > P)}$.
	Define $\ar_\text{linear}$ as the text “We agree that when the money amounts involved are large and the game is not repeated, there is no good reason to consider only expected gains when choosing a lottery. But in our case, the money amounts are small and the game is repeated. Therefore, expected gain is the right indicator to consider for a systematic and sound decision strategy.”
	Define $+$ as the concatenation operator over strings.
	Define $\dg = \set{(\ar_0, \ar^r_\text{non-linear}, \ar_0 + \ar_\text{linear}) \suchthat r \in \R}$. This strategy sets up the argument $\ar_0 + \ar_\text{linear}$ as an answer to any argument of the form described in \cref{ex:twolottstrings}, arguing for non-linearity of utility of money.
	This defines a descriptive theory of depth $2$.
\end{example}

\subsection{Validity and falsification}
\label{sec:validity}
An empirical theory claims to be valid, in the following sense.
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: \prop \in \iprops$.
\end{definition}
To achieve this goal, it claims that it is able to produce decisive arguments for its supported propositions.
These arguments can vary according to the individual $i \in I$.
Given an individual, the theory claims that a decisive argument will be systematically found by starting from an initial argument given by $\gleadsto$, and following a chain of defense defined by $\dg$ as required, depending on the paths that challenge the arguments produced along the way, as judged by the individual $i$. The property that a theory manages to systematically produce decisive arguments in such a way is named Convincingness. Defining it precisely requires to define first the notion of required arguments.

Given $i \in I, \prop \in \gpropse$, define $\dgip(\ar) = \set{\dg(\ar, \hist) \suchthat \ar \notin \histend \land \hist \nileadsto \phi} \subseteq \dg(\ar)$, representing the subset of defenders of $\ar$ required for convincing $i$ of $\prop$.
Given any set $S$ and function $f: S → S$, let $f^0$ represent the identity function, and given $k \in \N$, let $f^{k + 1} = f \circ f^k$.
Define $\gargs \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\prop)$ under $\dgip$: $\gargs = \cup_{k \in \N} (\dgip)^k({\gleadstoinv}(\prop))$. The set $\gargs$ is the set of arguments required to convince $i$ of $\phi$.
\begin{example}[Required arguments]
	\label{ex:required}
	Continuing \cref{ex:infinite}, define the subset of individuals that find some argument $\ar^r_\text{non-linear}$ convincing in presence of $\ar_0$ as the set of $i \in I$ such that $\exists \hist \in \allhist \suchthat \ar_0 \in \histend \land \ar^r_\text{non-linear} \in \range(\hist) \land \hist \nileadsto D > P$.
	We see that for any $i$ in that set of individuals, $\gargs[D > P] = \set{\ar_0, \ar_0 + \ar_\text{linear}}$, and for any $i$ in the complementary set of individuals, $\gargs[D > P] = \set{\ar_0}$.
\end{example}

Given $\ar \in \gargs$, say that $\ar$ $i$-defends $\phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend \land \hist \nileadsto \phi: d(\ar, \hist) ≠ \emptyset$.
\begin{definition}[Convincingness]
	An empirical theory is \emph{convincing} iff $\forall i \in I, \prop \in \gpropse, \ar \in \gargs: \ar$ $i$-defends $\prop$.
\end{definition}
\begin{example}[Convincingness]
	Pursuing with \cref{ex:required}, that theory will be convincing iff either $i$ finds some argument $\ar^r_\text{non-linear}$ convincing in presence of $\ar_0$ and both $\ar_0$ and $\ar_0 + \ar_\text{linear}$ $i$-defend $D > P$, or $i$ does not find any argument $\ar^r_\text{non-linear}$ convincing in presence of $\ar_0$ and $\ar_0$ $i$-defends $D > P$.

	$\ar_0$ $i$-defends $D > P$ iff any argumentative path $\hist$ that contains $\ar_0$ towards its end and that does not lead $i$ to accept $D > P$ is such that $\dg(\ar_0, \hist) ≠ \emptyset$, or equivalently, is such that the theory has some argument to answer to $\hist$, or still equivalently, is such that it contains some $\ar^r_\text{non-linear}$.
	This property mandates that no argumentative path has been “forgotten” in $\dg$: if some path is found that attacks $\ar_0$ successfully (thus $\ar_0 \in \histend \land \hist \nileadsto D > P$), and if the theory has nothing to answer to it (thus $\dg(\ar_0, \hist) = \emptyset$), then the theory is not Convincing.
\end{example}

It is now necessary to verify that the claim of convincingness is, first, falsifiable, and second, sufficient to yield validity.

Let us first clarify what falsifiability formally means in our context. So far, convincingness is defined only for the “real” relation $\leadsto$; we now need it defined more generally. Given any relation $R \subseteq I × \allhist × \Phi$, say that a descriptive theory is convincing under $R$ whenever it is convincing, as defined above, but replacing $\leadsto$ by $R$. 
We can now speak counterfactually of the theory being convincing, or not, assuming the observable relation is $R$ instead of $\leadsto$. 
Observations of the relation $\leadsto$ can in general include positive statements, of the form $\hist \ileadsto \phi$, and negative ones, of the form $\hist \nileadsto \phi$. (One way to obtain such negative observations is to use an incompatible proposition $\phi'$ and deduce from $\hist \ileadsto \phi'$ that $\hist \nileadsto \phi$.) A relation $R \subseteq I × \allhist × \Phi$ is compatible with the positive observations $O \subseteq I × \allhist × \Phi$ and the negative observations $O' \subseteq I × \allhist × \Phi$ iff $O \subseteq R \land R \cap O' = \emptyset$.
The claim of convincingness is said to be falsifiable iff there exists finite set of observations $O, O' \subseteq I × \allhist × \Phi$ such that $\forall R \subseteq I × \allhist × \Phi \suchthat O \supseteq R \land R \cap O' = \emptyset$, the theory is not convincing under $R$. 
In such a case, no possible relation $R$ compatible with the observations $O, O'$ make the theory convincing; in other words, the observations $O, O'$ prove that the theory is not convincing.
%\begin{example}[Two claims]
%	Given a normative theory and $\phi \in \Phi$, the claim $C_1 = \set{{\leadsto} \subseteq I × \allhist × \Phi \suchthat \forall i \in I, \exists \ar \in \allargs \suchthat \ar \ileadstost \phi}$ corresponds to claiming that for all individuals $i \in I$, some argument is $i$-decisive for $\phi$. If the set of arguments is infinite, $C_1$ is not falsifiable.  
%	Given $\ar \in \allargs$ and $\phi \in \Phi$, the set $C_2 = \set{{\leadsto} \subseteq I × \allhist × \Phi \suchthat \forall i \in I, \ar \ileadstost \phi}$ denotes the claim that all individuals will consider the argument $\ar$ as decisive for $\phi$. This claim is falsified by any observation such that $\hist \nileadsto \phi$ for some $\hist \in \allhist$ with $\ar \in \histend$, thus, $C_2$ is falsifiable.
%\end{example}

The following theorem states that the descriptive theories defined here are falsifiable, unless they state nothing about anybody’s \ac{DJ}. \Cref{sec:prooffals} proves it and describes the set of falsifiers of such theories.
This theorem implicitly assumes full observability, that is, it supposes that any positive or negative observations are in principle observable and thus available for falsification attempts. Further works should investigate to which extent this hypothesis can be relaxed.
\begin{theorem}
	\label{th:convfals}
	The claim of convincingness is falsifiable iff ${\gleadsto} ≠ \emptyset$.
\end{theorem}
It can also be seen that if a theory is not constrained in its depth, its claim of convincingness may be unfalsifiable, thus the hypothesis of finite depth is required to obtain this theorem. The formal proof of this fact is omitted because it is not especially enlightening; that this fact holds can be seen by examining the use of this hypothesis in the proof of \cref{th:convfals}.

Only remains now to prove that convincingness, together with normative adequacy, is sufficient to obtain validity. The proof is also provided in \cref{sec:prooffals}.
\begin{theorem}
	\label{th:convvalid}
	A normatively adequate empirical theory that is convincing is valid.
\end{theorem}

\subsection{Effectively comparing models}
\label{sec:compare}
Recall from \cref{sec:coh} that $\incompat = \set{(\prop, \prop') \suchthat \forall \hist \in \allhist: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'}$ denotes the set of pairs of incompatible propositions.
\commentOC{Does this mandate completeness?}
Given two empirical theories $\gamma$ and $\delta$ for a given normative theory, say that they disagree about $(\prop, \propbar) \in \incompat$ iff $\prop \in \gpropse[\gamma] \land \propbar \in \gpropse[\delta]$. Say that the theories are incompatible iff there is at least one pair $(\prop, \propbar) \in \incompat$ that they disagree about.

A path $\histp$ starts with a path $\hist$ iff $\card{\histp} ≥ \card{\hist}$ and $\forall k \in \intvl{1, \card{\hist}}: \histp_k = \hist_k$.
An increasing sequence of paths is a finite sequence $({\hist}^k)_{k \in \intvl{1, N}}$, for some $N \in \N$, of paths, 
such that $\forall k \in \intvl{2, N}, {\hist}^k$ starts with ${\hist}^{k-1}$.

\begin{theorem}
	\label{th:proc}
	Given two incompatible empirical theories for a given normative theory, $\gamma = (\gleadsto[\gamma], \gbeats[\gamma], \alldargs_\gamma, \dbeats[\gamma], +_\gamma)$ and $\delta = (\gleadsto[\delta], \gbeats[\delta], \alldargs_\delta, \dbeats[\delta], +_\delta)$ and given any $i \in I$, there exists a procedure that falsifies one of these theories, queries only $i$, and queries $i$ only about some increasing sequence of paths.
\end{theorem}
That procedure is described in \cref{sec:proc}.

\section{Multiple}
\label{sec:multiple}

\appendix
\section{Old introduction} 
Decision theoretic approaches in economy are classically divided into normative and descriptive approaches. Normative approaches aim at prescribing decision behavior; descriptive approaches aim at describing decision behavior. These two goals are generally considered incompatible: what is normative is not to be judged on descriptive grounds, and conversely. (Some authors consider prescription as the art of using descriptive insights to inform recommendation \citep{bell_decision_1988}, a view that also considers the normative aim as distinct to the descriptive aim.)

As a result, normative approaches are difficult to validate: being considered as non-descriptive in their goal, normative claims cannot be submitted to empirical tests that would permit to verify or falsify them. Observing a behavior that deviates from a prescribed norm can be taken as a sign that the norm is useful: a decision maker does not hire analysts just to be prescribed the choice that she would have naturally opted for.
Instead, the validity of a normative approach is generally considered to be determined by examining the convincingness power of the normative axioms on which it rests.
\commentYM{par qui ? dans le cadre de l'analyse ou en amont ? et pourquoi sur la seule base de la soundness de l'axiome ? pourquoi pas sur la base de ses implications dans des cas concrets ?}
This validity criterion is unsatisfactory, as it is unable to settle disputes between experts when they arise. In supplement, agreement between experts in decision theory is arguably not enough to ground the practical applicability of some normative axioms: one must ensure that decision makers to which the prescription is addressed agree with the particular instanciation of these axioms in the specific decision problem under study. But no systematic, formally defined procedure has been proposed to proceed with this check, to the best of my knowledge. 
In practical applications, an expert may informally discuss the axioms with the decision maker in order to check endorsement, either in general, or by discussing violations on a case by case basis when they are observed during an interactive elicitation session. 
The resulting state of conviction arguably depends in part on the rethorical ability of the expert, not solely on the normative strength of the axioms. It may lead to a decision maker accepting a decision procedure considered unsound by many experts in the field, such as the \ac{AHP} \citep{smith_anniversary_2004}. 
This problem is compounded by the enormous constraining power gained when axioms interact, as the celebrated theorem of Arrow shows dramatically: axioms that appear innocuous when taken alone may precipitate undesirable results when combined.

By contrast, this difficulty of validation is much alleviated with descriptive approaches. Such approaches can be validated by confronting their predictions to empirical data. When two descriptive theories disagree on their predictions, it is in principle possible to show that at least one of them is wrong with a suitable experiment. 
\commentYM{quelle différence avec les implications empiriques d'une théorie normative ?}
\commentOC{Est-ce plus clair maintenant ?}
Alàs, descriptive approaches as currently existing in decision theory are not suitable to obtain valid prescriptions, as we do not merely want to prescribe natural, unreflectful behavior.

This article defines decision theories, called theories of \ac{DJ}, that are both normative and descriptive: they claim to be able to yield valid prescriptions, and these prescriptions are claimed to predict reflective behavior. Reflective behavior as defined in this article is observable, by means of observing the reaction of individuals to arguments. This allows to confront theories of \ac{DJ} to empirical data.

This article defines theories of \ac{DJ} in general, not a specific theory of \ac{DJ}. Specific theories of \ac{DJ} are tailored to a set of situations and yield claims relevant to those situations. For example, a specific theory of \ac{DJ} may be tailored to situations of choice between lotteries involving money. This article adopts a more general point of view. It defines a general form for theories of \ac{DJ}, and proves claims that hold for any specific theory of \ac{DJ}, showing that such theories are indeed normative and descriptive in some technical sense to be defined.

Here is an example opposing two simplistic theories of \ac{DJ}, meant to give a flavor of the sort of theories this article aims at defining.

\begin{example}[Two disagreeing theories]
	\label{ex:two}
	Consider a set $\allalts$ of items evaluated on two criteria, say, price and speed, using functions $g_1$ and $g_2$, where $g_i: \allalts → \R$ measures the performance of items on the criterion $i$, and a set of individuals $I$ whose reflective choice among $\allalts$ we are interested in. 
	Consider the following the theories of \ac{DJ}.

	The \emph{satisfycing} theory claims that any individual $i \in I$ ought to choose, and will reflectively choose, whichever item is satisfying (and be indifferent if they are both satisfying or if none of them is). An item $\alt$ is satisfying iff $g_1(\alt) < t_1 \land g_2(\alt) > t_2$, where $t_1$ and $t_2$ are two real numbers determined by the theorist.

	The \emph{additive} theory claims that any individual $i \in I$ ought to choose, and will reflectively choose, whichever item $\alt$ maximises $v(\alt) = v_1(g_1(\alt)) + v_2(g_2(\alt))$, where $v_i: \R → \R$ are functions  chosen by the theorist.

	Consider a set $\alts = \set{\alt_1, \alt_2} \subset \allalts$ of two items, and assume the two theories of \ac{DJ} are opposed on their prediction of the reflective choice of individuals having to choose one item among $\alts$. Say, $\alt_1$ is satisfying and $\alt_2$ is not, failing on the speed criterion; while $v(\alt_2) > v(\alt_1)$.

	To determine which theory is right, it is insufficient to simply observe what an individual $i \in I$ effectively chooses. This would merely represent her unreflective behavior. The approach adopted in this article to capture \acp{DJ} is to declare that the winning theory (if any) is the one that exhibit arguments able to systematically convince individuals that their proposed choice is to be adopted.
	The satisfycing theory could argue as follows: “$\alt_1$ is better than $\alt_2$, because it has a satisfactory price (lower than $t_1$, an acceptable price) and a satisfactory speed (higher than $t_2$, an acceptable performance); whereas $\alt_2$ is not satisfactory on the speed criterion”. 
	The additive theory could use the following argument: “$\alt_2$ is better than $\alt_1$, because $\alt_2$ is cheaper than $\alt_1$, and this more than compensates for its weaker performance in terms of speed”.

	The basic intuition on which this article is based is that determining which theory is right requires to present both arguments to $i$, then let her choose one item in $\alts$, to see which argument she finds more convincing.
\end{example}

The proposed approach starts with this intuition that the reaction of individuals to arguments should matter, and presents a general solution to the problem of capturing reflective judgments or choices. Contrary to the simplistic situation of \cref{ex:two}, this solution must allow to confront more than just two arguments or two theories. It should define precisely what is required for a judgment to be called deliberated, and determine formally which hypothesis are required for the prescriptions of a theory of \ac{DJ} to be valid.

\section{Normative theories (focused)}
Define a normative axiom as a formal constraint on behavior that constrains behavior so as to yield valid prescription. 
Define a descriptive axiom as a formal constraint on behavior interpreted as an empirical prediction, that is, as a claim that behavior (in some a priori described circumstances) satisfies the constraint.

A normative axiom has to be accepted on the basis of its apparent adequacy to capture what is intuitively an acceptable requirement for a prescription to a \ac{DM}. In other words, it is not a hypothesis on the observable world: that the requirement holds cannot be proven by observation. Examples are given by the axioms of classical decision theories, when they are applied for prescription. Consider for example Savage’s sure thing principle. One can observe whether individuals unreflectively behave according to Savage’s sure thing principle, but even when they fail to, it can be said that this is because of a mistake: the normative status of an axiom is not hit by violations due to unreflective behavior. Such axioms are not falsifiable by behavioral observation. Conversely, observing that some axiom perfectly matches observed behavior, that empirical observation alone does not yield any validity to prescription based on that constraint: accepting the validity of a prescription rests on accepting the soundness of its reasoning, not on its ability to reflect natural behavior.

Let us assume that some natural behavior is not to be prescribed.

At least one such axiom is required to ground any prescription. Hence, no theory can be fully descriptive, and theories of \ac{DJ} make no exception. The proposed path out of this conundrum is to restrict normative axioms to a self-evident claim, and leave most of the constraining power to descriptive claims.

The pursued path is as follows.
This article introduces first a normative axiom that mandates that whatever propositions the \ac{DM} accepts when reasoning carefully can be prescribed. This notion of careful acceptability is defined formally on the basis of an observable relation that captures the reactions of the \ac{DM} to arguments.

Second, this article defines descriptive theories of acceptable propositions. Acceptable here is in the sense of the careful acceptability just defined informally. Descriptive here means that these theories can predict at least some of the propositions that will be considered acceptable by the \ac{DM}.
This is arguably the more important and difficult part of the current research program: it has to convince that it is practically possible to define theories that predict such an a priori volatile thing as the way individuals reason. It is plain that mandating that theories be able to predict the reaction of individuals to every possible sequence of arguments they are exposed to would be plainly irrealistic. Some weaker criterion is required. How weak can it become? As announced, the proposed theories must be purely descriptive, apart from the required acceptance of the normative adequacy axiom described in the above paragraph. In other words, their conclusions must hold on the sole basis of the unique normative axiom, aided by claims that can be proven false experimentally. Equivalently, if the theory is invalid, in the sense of failing its predictions about the acceptability of propositions, then it must be falsifiable: there must exist a set of observable reactions of the \ac{DM} to arguments that contradict the claims of the theory. Accordingly, this article claims and proves that empirical theories of \ac{DJ} as defined here are such that under normative adequacy, if the theory is not falsified, then it is valid. I admit readily that this might not suffice to convince that building such concrete theories is feasible. It is only one of the difficulties to tackle.

A possible objection to the usefulness of this approach should perhaps be lifted before delving into the formal descriptions of this rough sketch.
It may be argued that because this proposition needs a normative axiom, which cannot be validated empirically, it does not fare better than the alternative propositions existing in the literature. 
%\ac{SEU}, for example, also can be considered to claim to obtain reflective preferences on the basis of normative axioms.
I grant that it is impossible to get completely rid of normative axioms, because prescriptions must be linked to descriptions, and this cannot be done in a complete empirical way.
However, it is good to lighten as much as possible the content of normative axioms, and defer as much content as possible to descriptive claims. The descriptive parts of theories can be confronted to observations, a natural way to lift disagreements that has often proved useful in the history of science.
I postulate that the normative axiom proposed here will be consensual, at least for some non trivial concrete instanciations of the proposed approach, among economists, practitioners, external observers and \acp{DM}. Furthermore, its simplicity and a priori obviousness are strong points in its favor. It has, in some intuitive sense that cannot be described formally, less “content”, in the sense of that word used when economists argue that their theories should not have content about the preferences of the individuals (for fear of being paternalistic or dictatorial). Of course, any decision theory has content in a wider sense of imposing some structure to preferences, otherwise, the theory is void. The question is: are the restrictions imposed of a kind that can be judged acceptable with respect to the considered applications of the theory? Favoring judgments that resist counter-arguments, as done here, seems to me to capture exactly what is desired in a prescription context.
%Arguably, the normative axiom proposed here is much weaker than those of a theory such as \ac{SEU}, in the sense of having less substance. (This claim can only be valid at an intuitive level: because \ac{SEU} does not use arguments, its axioms cannot be compared formally to mine.) 
The lack of concrete substance in this normative axiom does not prevent theories of \ac{DJ} to propose substantive, non-trivial claims about reflective preferences. It simply moves the debate from the world of discussions between experts about the validity of the axioms to the world of observations of acceptability of arguments by the persons to which these theories claim to be applicable.
It is also worth noting that the proposed axiom has a specific form that may permit \acp{DM} to reconstruct the reasoning proposed by the theory so as to be able to check that they apply to their situation. Detailed exploration of this aspect is deferred to future work.

The rest of this article is structured as follows.
\Cref{sec:approaches} defines more precisely what is meant here with normative and descriptive theories. These definitions permit to readily see why the apparent impossibility of a claim being both normative and descriptive is an illusion, thereby enabling a solution to what could be called the “normative against descriptive” gap. It also reviews related works in the literature, and briefly sketches the current proposal.
\Cref{sec:normative,sec:descriptive} define the theories of \ac{DJ} and prove that these theories are normative and descriptive in the sense defined in \cref{sec:approaches}. \Cref{sec:preferences} define theories of \ac{DP} on the basis of theories of \ac{DJ}, and \cref{sec:conclusion} concludes.

\section{Eliciting reflective preferences}
\label{sec:approaches}
\subsection{Species of theories}
For this article, a \emph{normative decision theory} is a theory that, when applied in the situations it self-describes as applicable to, claims to yield valid prescriptions. I consider a prescription in the wide sense: recommending against some behavior, or recommending conditionally, also counts as prescribing.

A \emph{descriptive decision theory} is a theory that, when correctly applied in the situations it self-describes as applicable to, claims to predict the decision behavior of a given set of individuals. For this article, a claim is a prediction only when it is in principle possible to show that it is false using empirical data.

I will simply write normative theory and descriptive theory from now on to mean normative and descriptive decision theories.

For example, the \ac{SEU} theory of Savage is a normative theory. It may be used to recommend conditionally some behavior: “if you choose this lottery rather than that lottery, then you ought to \emph{not} choose this other lottery rather than that other lottery”.
Examples of descriptive theories include the many variants of Prospect theory.

These definitions avoid to define the concept of theory itself, for defining precisely such a fundamental concept raises serious difficulties. Is Prospect theory in general a theory, or only some of its variants, sufficiently precisely defined? Can a theory be extremely specific, or is some level of generality required to deserve being called a theory? Must a theory consist in a non-trivial assembly of laws, some deducible from others? It is not necessary for the purpose of the present discussion to take position on these issues, rather, the focus is on the distinction between normative and descriptive approaches. An intuitive understanding of the concept of theory will suffice.
Similarly, I sidestep insuperable difficulties pertaining to capturing in a short definition the concept of a descriptive theory in general: some descriptive theories in a more general sense may propose to describe without predicting, for example (historical decisions or trends come to mind); such endeavor simply does not enter the scope of this discussion. For similar reasons, I do not claim that the term normative is unanimously used in the literature as defined here. The present definition encompasses theories used in approaches that do not necessarily self-describe as normative \citep{bell_decision_1988, meinard_rationality_2019}; conversely, some theories considered as normative in the literature, such as some social choice theories, may not fall into the definition proposed here, depending on how they are interpreted.

I say that a normative theory is also a descriptive theory when the prescriptions of the normative theory are also claimed to predict the decision behavior of a given set of individuals.

Another contrast can now be introduced: the contrast between shallow and reflective behavior. \emph{Shallow behavior} is defined as the everyday, usual behavior of an individual. 
This is the behavior whose observation defines what \citet[p.\ 16]{von_neumann_theory_2004} describe as the “intuitive” preferences of an individual, those originating from an “immediate sensation” (the quotes from their article used here are cited by \citet{fishburn_retrospective_1989}).
\emph{Reflective behavior}, by contrast, is defined as the behavior that enacts informed, pondered decisions, those that would be taken after due consideration.
% an individual has consciously adopted after having pondered everything that is possibly relevant to the decision. 
%This pair of definitions is, again, vague and unsatisfactory for a deep philosophical discussion, but hopefully sufficient for the present one. 
%What is meant by “pondering about everything that is possibly relevant to the decision” will be defined more precisely later; an intuitive understanding of this phrase will do at this stage.
This distinction draws on our capacity, as human beings, to think about our shallow preferences and consider some of them as mistaken \citep{frankfurt_freedom_1971}. 
A famous example of such a situation is the one of \citet[pp.\ 101--103]{savage_foundations_1972} having realized, after due consideration, that his initial declared preference in a decision problem presented by \citet{allais_so-called_1979} was mistaken.

This contrast raises a new distinction, which appears by simple combination of the just given definitions: a \emph{shallow descriptive theory} is a descriptive theory that claims to predict the shallow behavior of individuals; a \emph{reflective descriptive theory} is one that claims to predict the reflective behavior of individuals.

The descriptive theories proposed in decision theory, in the sense defined here, are mostly shallow descriptive theories, to the best of my knowledge (\cref{sec:purification} discusses the particular case of “preference purification” approaches).
%This need not be so. Indeed, this article introduces descriptive theories about reflective behavior.
This observation permits to understand what could be called the “normative versus descriptive gap”, referring to the belief, implicit in the decision theoretic literature, that a theory cannot be both normative and descriptive. Because normative theories aim to prescribe, they are required to not yield recommendations that precisely, systematically match shallow behavior. Given the numerous results in experimental psychology showing that shallow behavior is undefensible in many situations, a normative theory whose prescriptions would systematically match shallow behavior would be suspicious, to say the least. It would also be of low practical interest.
It is an immediate consequence of this minimal requirement that a normative theory cannot also be a shallow descriptive theory. 
In other words, as \citet{hausman_preference_1994} illustrate with an example of a person drinking a fatal poison while erroneously thinking that it is water, shallow behavior is not sufficient to inform normative economics.
The “normative versus descriptive gap” comes from an implicit identification of descriptive theories with shallow descriptive theories.

This gap can thus be overcome by considering reflective descriptive theories, instead of shallow descriptive theories. This is the path this article adopts.

%This is done by defining on the one hand a non-observable set, containing those propositions that are valid prescriptions; and on the other hand an observable set, containing those propositions decisively defended by arguments, as judged by a given individual. This latter set is called the individual’s \ac{DJ}.
%A normative axiom, that any theory of \ac{DJ} rests on, links these two sets by stating that what is in someone’s \ac{DJ} is a valid prescription.
%This justifies the normative aspect of the theories. 
%The theory yields descriptive claims thanks to the

\subsection{The proposed approach}
The proposal elaborated in this article defines theories that are both normative, and descriptive of reflective behavior.
The kind of behavior of interest here is choice behavior, although the proposal is more general.

In fact, the main object of study of this article is judgments about propositions. The main goal is accordingly to study reflective adherence to propositions.
This will then be applied to determining reflective choice behavior. Choice behavior is obtained when considering adherence to propositions of the form: “I ought to choose alternative $\alt$ rather than $\alt'$ if both are available”. The alternatives can be concrete objects such as consumable goods.

One of our main preoccupation will be to clarify, using formal and precise definitions, the intuitive notion of reflective adherence to propositions. This is done using the concept of \emph{deliberated} judgments (a normative concept) and of \emph{acceptable} propositions (a descriptive concept).

A theory of \ac{DJ} applies to a set of individuals and a set of propositions. The \ac{DJ} of an individual $i$ is the set of propositions $\prop$ such that it is legitimate to recommend to $i$ adherence to $\prop$. Someone’s \ac{DJ} is not directly observable.
However, we can observe the reactions of $i$ to arguments in favor and against the propositions.
A theory of \ac{DJ} assumes that it knows a set of arguments such that exploring the reactions to these arguments “sufficiently” permits to obtain knowledge about someone’s \ac{DJ}.
Thus, one important task dealt with in \cref{sec:normative} will be to define the notion of acceptability, that serves to capture formally this notion of “sufficient” exploration. Intuitively, a proposition is acceptable when some argument makes $i$ seem (according to the observations) to adhere to the proposition whatever the counter-arguments given to $i$, and whatever the sequence of arguments given to $i$ previously.

A theory of \ac{DJ} must use an axiom that relates the descriptive notion of acceptability to the normative notion of being deliberated. This axiom simply mandates that whatever is acceptable, is deliberated. \Cref{sec:normative} presents it formally.

This defines theories of \ac{DJ} that can be considered normative, as accepting the axiom is sufficient to determine unambiguously a set of legitimate prescriptions to be given to individuals. 
This is done by linking \acp{DJ} to acceptable propositions, a notion whose major interest is that it can be confronted to empirical observations. But at this stage, it is still unclear how to obtain knowledge of someone’s \ac{DJ}, and accordingly, such theories cannot still be considered as descriptive. In particular, they are not falsifiable, as they take no position on the reactions of individuals to arguments. \Cref{sec:descriptive} accordingly defines descriptive theories of acceptable propositions, assuming given a normative theory of \ac{DJ}.

Such descriptive theories can take multiple forms, and some of them would permit to see clearly that the theory is descriptive in the required sense, that is, that its claims can be falsified given observational data. 
But the task we face, for such theories to be useful, is more complex than this minimal requirement: we have to find forms of theories that are not hugely demanding on the part of the theorist. In particular, we do not want to oblige the theorist to predict the whole of the argumentative attitude of individuals if this is not required to obtain information about her \ac{DJ}. Thus, \cref{sec:descriptive} defines descriptive theories of \ac{DJ} that only claim to be able to answer counter-arguments, in a sense that will be made precise. The desire that its claim be falsifiable is also defined formally. More formally, the result that is obtained in that section, and that is the central result of this article, is that if a theory resists all empirical falsification attempts, then it is valid, assuming acceptance of the normative axiom described here above informally (and formally in \cref{sec:normative}). A theory is valid if the propositions that it claims are in someone’s \ac{DJ} are indeed in that person’s \ac{DJ}. This central result thus amounts to prove the main claim of this article, namely, that accepting a weak normative axiom, supplemented with empirical claims, is enough to capture someone’s reflective preferences.

One long-term goal of this research program is to give individuals the means to reject theories that they find unconvincing after due consideration of all arguments; and accordingly, give incentives to decision theorists to build defenses of these theories able to convince individuals of the soundness of these theories. \commentYM{phrase trop longue}This might match \possessivecite{hausman_debate_2010} suggestion that we should use rational persuasion rather than nudge individuals, arguing that it would promote autonomy.
It also follows \possessivecite[p.\ 72]{guala_logic_2000} suggestion (discussing the Allais paradox) about obtaining relevant data to compare normative behavioral theories: “Before asking whether agents want to revise their choices or not, \emph{both} rival theories should be proposed” (emphasis in the original text).

Let us now examine how the approach proposed here to capture reflective preferences differs from those proposed in the literature.

The next two sections review existing approaches aiming at obtaining a model of the reflective preferences of an individual. They distinguish two kinds of approaches. \Cref{sec:purification} mentions formal techniques, that can be applied by following an explicit procedure, but that do not consider the reaction of individuals to arguments; while \cref{sec:informal} reviews studies that involve informal discussions with individuals, which do consider arguments, but require involving experts. By contrast, this article presents theories that can be confronted to empirical data in a formal, systematic way, and that consider the reactions of individuals to arguments.

Before delving into these two categories of approaches, let me mention a few others that, though related to the topic discussed here, fall outside the scope of this article.

Some authors have proposed to base normative theories on observable aspects of nature and individuals; for example, \citet{kahneman_back_1997} propose to measure “experienced utility”, considered as a hedonic quality. 
To the best of my knowledge, these proposals never aim at observing reflective preferences as considered here and thus fall outside the scope of this article.

The field of formal argumentation theory has developed technical approaches to deal with arguments. Such endeavor starts with an “attack” relation, which indicates which arguments argue against which. Various criteria can then be mathematically defined, in the form of a function selecting subsets of arguments given an attack relation, typically representing sets of arguments that “best resist” their counter-arguments. This field has greatly inspired this proposal, in particular, the notion of attack and of reinstatement. This field however does not view the attack relation as observable, but considers it as given, and its methodology is therefore not empirical in the sense seeked here. (\Citet{cailloux_formal_2019} elaborates somewhat on this difference of objective.) 
\commentYM{ici aussi, je trouve la mention de ces approches maladroite, car elles sont mises en exergue comme si elles étaient des manières de traiter le problème qui t'intéresse, ce qu'elles ne sont pas}

The field of Empirical social choice \citep{gaertner_empirical_2012} investigates individuals’s judgments about fairness: in a typical laboratory experiment, subjects are described some setting requiring to distribute resources among a set of individuals, and are asked to choose a distribution that they consider fair. These interesting approaches permit to study empirically the sense of fairness of real individuals, and compare them to social choice theories. They are distinct from the currently proposed approach as they seek to obtain their a priori sense of fairness, rather than aim at observing how subjects would appreciate (possibly counter-intuitive) claims about justice when they are backed by (possibly elaborate) arguments; and do not specifically aim to prescribe to an individual.

The phrase “\acl{DP}” has been used by \citet{orchard-webb_deliberative_2016, kenter_deliberative_2017} in the context of group deliberation. As used in this article, the concept of deliberation does not necessarily refer to a group deliberation: this approach is agnostic as to the provenance of the arguments.

>>>>

\hbadness=10000
\bibliography{simple}
\end{document}
